\propertitle

### 2019-06-18 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-18T19:00:36+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
% https://tex.stackexchange.com/a/33547/97039
%\newcommand*{\fm}{f}
\newcommand*{\px}{p_{\text{me}}}

@@

These two steps are more straightforward if instead of power moments we use
\emph{normalized factorial moments} \citep{potts1953}. The $m$th normalized
factorial moment of a distribution $\px(\ya)$ for the activity of the
sample neurons is defined as the average
\begin{equation}
  % \E\condt[\Bigg]{
  %   \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  % \equiv
  \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \px(\ya),
  \qquad 1\le m \le n
\end{equation}
% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
% new bin.
This moment can be interpreted as the expectation of the number of distinct
$m$-tuples of simultaneously spiking neurons (within a bin's time width),
normalized by the number of distinct $m$-tuples. For example, with $m=2$,
if $na=4$ neurons spike in a network of $n=5$, we have $\tbinom{4}{2}=6$
distinct pairs of simultaneously spiking neurons, and the total number of
distinct pairs is $\tbinom{5}{2}=10$. The normalized number of spiking
pairs is therefore $6/10$. Note that the first $m$ factorial moments
linearly determine the first $m$ power moments and vice versa, because
$\tbinom{n\ya}{m}$ is a polynomial in $\ya$ of degree $m$; so fixing the
ones is equivalent to fixing the others. Now we have this extremely
convenient property: \emph{the first $n$ normalized factorial moments for a
  sample and for the full network are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
    \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \px(\ya)
=    \sum_{\yA} %m!\,
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA),
  % \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  % \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  \qquad 1\le m \le n,
\end{equation}
where $\px(\yA)$ is the distribution for the full-network activity.

We can therefore apply the maximum-entropy method to obtain a distribution
for the full network, by constraining its $m$th factorial moment to be
equal to the sample's recorded average of
$\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$, for as many $m$ as we please with
$1\le m \le n$. The constraints on the distribution $\px(\yA)$ for
the full network are
\begin{equation}
  \label{eq:final_constraint}
  \frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
  =
   \sum_{\yA}
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA)
\end{equation}
for any $m$s we wish.

@@

The particular case in which $n$ moments are constrained is especially
important: it corresponds to fully specify the measured frequency
distribution for the activity of the sample neurons. In this case, our
inferences about the full-network activity are using all available measured
frequency data. Note that application of the maximum-entropy method
\emph{at the sample level} is in this case trivial and meaningless -- it
just gives back the measured frequency distribution. But application of the
method \emph{at the level of the full network}, as done here, is not
trivial.

Using the full frequency distribution of the sample may be a bad idea,
however, because the maximum-entropy distribution becomes a bad
approximation of some of the four distributions described above. It is
preferable to use a moderately high number of moments smaller than $n$. We
explain this point in \sect***.

@@

In the next section we apply the method to the data sets from two actual
recordings, using *** moments, and discuss the properties of the resulting
distributions.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-18T19:09:53+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
% https://tex.stackexchange.com/a/33547/97039
%\newcommand*{\fm}{f}
\newcommand*{\px}{p_{\text{me}}}
\newcommand*{\pxx}{p}

@@

These two steps are more straightforward if instead of power moments we use
\emph{normalized factorial moments} \citep{potts1953}. The $m$th normalized
factorial moment of a distribution $\pxx(\ya)$ for the activity of the
sample neurons is defined as the average
\begin{equation}
  % \E\condt[\Bigg]{
  %   \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  % \equiv
  \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pxx(\ya),
  \qquad 1\le m \le n
\end{equation}
% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
% new bin.
This moment can be interpreted as the expectation of the number of distinct
$m$-tuples of simultaneously spiking neurons (within a bin's time width),
normalized by the number of distinct $m$-tuples. For example, with $m=2$,
if $na=4$ neurons spike in a network of $n=5$, we have $\tbinom{4}{2}=6$
distinct pairs of simultaneously spiking neurons, and the total number of
distinct pairs is $\tbinom{5}{2}=10$. The normalized number of spiking
pairs is therefore $6/10$. Note that the first $m$ factorial moments
linearly determine the first $m$ power moments and vice versa, because
$\tbinom{n\ya}{m}$ is a polynomial in $\ya$ of degree $m$; so fixing the
ones is equivalent to fixing the others. Now we have this extremely
convenient property: \emph{the first $n$ normalized factorial moments for a
  sample and for the full network are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
    \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pxx(\ya)
=    \sum_{\yA} %m!\,
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pxx(\yA),
  % \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  % \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  \qquad 1\le m \le n,
\end{equation}
where $\pxx(\yA)$ is the distribution for the full-network activity.

We can therefore apply the maximum-entropy method to obtain a distribution
for the full network, by constraining its $m$th factorial moment to be
equal to the sample's recorded average of
$\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$, for as many $m$ as we please with
$1\le m \le n$. The constraints on the maximum-entropy distribution
$\px(\yA)$ for the full network are
\begin{equation}
  \label{eq:final_constraint}
  \frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
  =
   \sum_{\yA}
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA)
\end{equation}
for any $m$s we wish.

@@

The particular case in which $n$ moments are constrained is especially
important: it corresponds to fully constraining the marginal frequency
distribution for the activity of the sample neurons. In this case, our
inferences about the full-network activity are using all available measured
frequency data. Note that application of the maximum-entropy method
\emph{at the sample level} is trivial and meaningless in this case -- it
just gives back the measured frequency distribution. But application of the
method \emph{at the level of the full network} is not trivial.

Using the full frequency distribution of the sample may be a bad idea,
however, because the maximum-entropy distribution may become a bad
approximation of some of the four distributions described above. It is
preferable to use a moderately high number of moments smaller than $n$. We
explain this point in \sect***.

@@

In the next section we apply the method just described to the data sets
from two actual recordings, using *** moments, and discuss the properties
of the resulting distributions.
